### 2017-09-27 paper.tex
Fixing DOI
Changing Greg's affiliation
Handling some of the smaller tasks as per #40 See checklist, not entirely sure about tweaks to Rule #3 and #5

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules for collaborative lesson development} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
{Gabriel~A.~Devenyi}\textsuperscript{1{\ddag}},
{R\'{e}mi~Emonet}\textsuperscript{2{\ddag}},
{Rayna~M.~Harris}\textsuperscript{3{\ddag}},
{Kate~L.~Hertweck}\textsuperscript{4{\ddag}},
{Damien~Irving}\textsuperscript{5{\ddag}},
{Ian~Milligan}\textsuperscript{6{\ddag}},
{Greg~Wilson}\textsuperscript{7{\ddag}*}
\\
\bigskip
\textbf{1} Douglas Mental Health University Institute, McGill University / gdevenyi@gmail.com \\
\textbf{2} Univ Lyon, UJM-Saint-Etienne, F-42023, France / remi.emonet@univ-st-etienne.fr \\
\textbf{3} The University of Texas at Austin / rayna.harris@utexas.edu \\
\textbf{4} The University of Texas at Tyler / khertweck@uttyler.edu \\
\textbf{5} CSIRO Oceans and Atmosphere / irving.damien@gmail.com \\
\textbf{6} University of Waterloo / i2millig@uwaterloo.ca \\
\textbf{7} Rangle.io / gvwilson@third-bit.com \\

@@

Despite the success of openness in software development and the curation of Wikipedia,
it is an uncommon approach in academic settings.
Each year,
thousands of university lecturers teach subjects ranging from first year biology,
to graduate-level courses in Indian film.
Some use textbooks written by one or a few authors,
but beyond that,
they develop and maintain their course materials in isolation.
This is curious given that research depends critically on sharing,
and that most researchers complain about how much time teaching takes away from research,
but the sociology and psychology behind this blind spot are out of the scope of this paper.

@@

The first requirement for building lessons together is
to know who they are being built for.
``Archaeology students'' is far too vague:
are you and your collaborators thinking of
first-year students who need an introduction to the field,
graduate students who intend to specialize in the sub-discipline which is the lesson's focus,
or someone in between?
If different contributors believe different things about prerequisite knowledge,
equipment or software required,
or how much time learners will have,
they will find it difficult to work together.

@@

Decades of pedagogical research has yielded many insights into
how best to build and deliver lessons \cite{hlw}.
Unfortunately,
many college and university faculty have little or no training in education \cite{brownell},
so this knowledge is rarely applied in the classroom.

@@

An example of how to teach these practices is Software Carpentry's instructor training program.
First offered in 2012,
is now a two-day course delivered both in-person and online
\cite{lessons-learned,instructor-training,how-to-teach-programming}.
This program teaches good pedagogical practices,
and introduces everyone who takes it to who Software Carpentry's lessons are for,
how they are delivered,
and how they are maintained.
Largely as a result of this training,
several hundred people per year now contribute to Software Carpentry's lessons.

@@

Complicating matters further,
some file formats make collaboration easier or more difficult.
Despite their ubiquity,
open source version control systems do not directly support review or merge
of Microsoft Office or LibreOffice file formats,
which raises an additional burden for newcomers \cite{jacobs}.
Programmers may look down on Google Docs and wikis
for their lack of pre-merge review and other capabilities,
but their low barrier to entry make them more welcoming to newcomers.

@@

A key part of doing this is to create opportunities for legitimate peripheral participation.
Curating a list of small tasks that newcomers can easily tackle,
encouraging them to give feedback on proposed changes,
or asking them to add new exercises and tweak diagrams and references
can all provide an on-ramp for people who might question their authority or ability to change the main body of a lesson.
Equally,
acknowledging all contributions,
however small,
gives new contributors an early reward for taking part.

@@

If lessons are being released regularly,
automate the process
and archive old versions in a discoverable location.
Also make sure that everyone involved knows what ``done'' look like,
i.e.,
which outstanding issues have to be addressed
and how it has to be formatted
in order for the next release to go out.
A simple checklist stored with the lesson materials is good enough to start,
but as time goes by,
the community may want to use an issue tracking system of some sort
so that work items can be assigned to specific people
and then ticked off as they are completed.

@@

Every day,
teachers all over the world spend countless hours duplicating each other's work.
These ten rules provide an alternative:
adopting the model of collaborative software development
to make more robust and sustainable lessons
that can be continually improved by those who use them.
We hope that our experiences can help others teach more
with more impact and less effort.

@@

\bibitem{jacobs}
Jacobs C.~T., Gorman G.~J., Rees H.~E., and Craig L.~E.:
Experiences With Efficient Methodologies for Teaching Computer Programming to Geoscientists
\newblock Journal of Geoscience Education
\newblock 64(3),
\newblock 2016,
\newblock https://doi.org/10.5408/15-101.1

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules for collaborative lesson development} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
{Gabriel~A.~Devenyi}\textsuperscript{1{\ddag}},
{R\'{e}mi~Emonet}\textsuperscript{2{\ddag}},
{Rayna~M.~Harris}\textsuperscript{3{\ddag}},
{Kate~L.~Hertweck}\textsuperscript{4{\ddag}},
{Damien~Irving}\textsuperscript{5{\ddag}},
{Ian~Milligan}\textsuperscript{6{\ddag}},
{Greg~Wilson}\textsuperscript{7{\ddag}*}
\\
\bigskip
\textbf{1} Douglas Mental Health University Institute, McGill University / gdevenyi@gmail.com \\
\textbf{2} Univ Lyon, UJM-Saint-Etienne, F-42023, France / remi.emonet@univ-st-etienne.fr \\
\textbf{3} The University of Texas at Austin / rayna.harris@utexas.edu \\
\textbf{4} The University of Texas at Tyler / khertweck@uttyler.edu \\
\textbf{5} CSIRO Oceans and Atmosphere / irving.damien@gmail.com \\
\textbf{6} University of Waterloo / i2millig@uwaterloo.ca \\
\textbf{7} DataCamp / gvwilson@third-bit.com \\

@@

Despite the success of openness in software development and the curation of Wikipedia,
it is an uncommon approach in the academic instructional setting.
Each year,
thousands of university lecturers teach subjects ranging from first year biology,
to graduate-level courses in Indian film.
Some use textbooks written by one or a few authors,
but beyond that,
they develop and maintain their course materials in isolation.
Given that academic research often depends on sharing, the differing
approach to developing pedagogical materials is interesting,
but the sociology and psychology behind this blind spot are out of the scope of this paper.

@@

The first requirement for building lessons together is
to know whom they are being built for.
``Archaeology students'' is far too vague:
are you and your collaborators thinking of
first-year students who need an introduction to the field,
graduate students who intend to specialize in the sub-discipline that is the lesson's focus,
or someone in between?
If different contributors believe different things about prerequisite knowledge,
equipment or software required,
or how much time learners will have,
they will find it difficult to work together.

@@

Decades of pedagogical research have yielded many insights into
how best to build and deliver lessons \cite{hlw}.
Unfortunately,
many college and university faculty have little or no formal training in education \cite{brownell},
so this knowledge is rarely applied in the classroom.

@@

An example of how to teach these practices is Software Carpentry's (\url{https://software-carpentry.org/}) instructor training program.
First offered in 2012,
is now a two-day course delivered both in-person and online
\cite{lessons-learned,instructor-training,how-to-teach-programming}.
This program teaches good pedagogical practices,
and introduces everyone who takes it to who Software Carpentry's lessons are for,
how they are delivered,
and how they are maintained.
Largely as a result of this training,
several hundred people per year now contribute to Software Carpentry's lessons.

@@

Complicating matters further,
some file formats make collaboration easier or more difficult.
Despite their ubiquity,
open source version control systems do not directly support review or merge
of Microsoft Office or LibreOffice file formats,
which raises an additional burden for newcomers \cite{jacobs}.
While Google Docs and wikis lack some capabilities, such as
full-fledged pre-merge review (although ``suggest mode'' 
mitigates this to some degree), 
their low barrier to entry makes them more welcoming to newcomers.

@@

A key part of doing this is to create opportunities for legitimate peripheral participation.
Curating a list of small tasks that newcomers can easily tackle,
encouraging them to give feedback on proposed changes,
or asking them to add new exercises and tweak diagrams and references
can all provide an on-ramp for people who might question their own authority or ability to change the main body of a lesson.
Equally,
acknowledging all contributions,
however small,
gives new contributors an early reward for taking part. 
It can also help with the formal and informal recognition
of labour, discussed in the following rule.

@@

If lessons are being released regularly,
automate the process
and archive old versions in a discoverable location.
Also make sure that everyone involved knows what ``done'' looks like,
i.e.,
which outstanding issues have to be addressed
and how they have to be formatted
in order for the next release to go out.
A simple checklist stored with the lesson materials is good enough to start,
but as time goes by,
the community may want to use an issue tracking system of some sort
so that work items can be assigned to specific people
and then ticked off as they are completed.

@@

Every day,
teachers all over the world spend countless hours duplicating each other's work.
These ten rules provide an alternative:
adopting the model of collaborative software development
to make more robust and sustainable lessons in all domains
that can be continually improved by those who use them.
We hope that our experiences can help others teach more
with more impact and less effort.

@@

\bibitem{jacobs}
Jacobs C.~T., Gorman G.~J., Rees H.~E., and Craig L.~E.:
Experiences With Efficient Methodologies for Teaching Computer Programming to Geoscientists
\newblock Journal of Geoscience Education
\newblock 64(3),
\newblock 2016,
\newblock 10.5408/15-101.1
